Detection of Oscillations in Aldebaran with Ground-Based Observations

### 2018-01-05 paper/aldebaran.tex
add comment back in
comments
Lots of technical CARMA crap, should eventually move most of it to Methods section.

@@

If we apply nonlinear statistical inference methods this situation can be improved. \citet{brewer2009} show that a system of driven, damped harmonic oscillators such as we encounter in asteroseismology can be statistically modelled as a Gaussian Process, with a covariance kernel consisting of a sum of damped sinusoids. The hyperparameters of this process can be used to retrieve very accurate power spectral density distributions. Unfortunately, this is impractical for long time series, as the computational cost of evaluating this likelihood function scales as $\mathcal{O}(n^3)$.

Reference \citet{Kelly2014}, compare to \citet{Foreman-Mackey2017}.

@@

%In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods. 
Aldebaran was observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27. Aldebaran and similar high-luminosity K giants are well known to show photometric variability due to solar-like oscillations \citep{bedding2000}, and hence the continuous, high-precision K2 light curves allow an independent confirmation of our CARMA results.

@@

\begin{figure}
\centering
% \includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
%\includegraphics[width=0.75\textwidth]{aldebaran_lc.eps}
\includegraphics[width=\textwidth]{insolation.png}
\caption{Evolution of the insolation of Aldebaran~b relative to the incident sunlight on Earth, as a function of the age of Aldebaran on the main sequence. The solid line is the mean of simulations, and the shaded region represents $1\,\sigma$ equivalent deviations. It is likely that Aldebaran~b spent some time on the main sequence with an equilibrium temperature similar to Earth.}
\label{insolation}
\end{figure}

>>>>

\newcommand{\dd}{\mathrm{d}}
\newcommand{\diff}[2]{\frac{\dd #1}{\dd #2}}

@@

If we apply nonlinear statistical inference methods this situation can be improved. \citet{brewer2009} show that a system of driven, damped harmonic oscillators such as we encounter in asteroseismology can be statistically modelled as a Gaussian Process, with a covariance kernel consisting of a sum of damped sinusoids. The hyperparameters of this process can be used to retrieve very accurate power spectral density distributions. Unfortunately, this is impractical for long time series, as the computational cost of evaluating the standard Gaussian process likelihood function scales as $\mathcal{O}(n^3)$.

A Gaussian process such as that used by \citet{brewer2009} also
corresponds to the solution, $y(t)$, of the stochastic ODE
\begin{equation}
  \label{eq:GP-definition}
  \prod_{i=1}^{p} \left[ \diff{}{t} - r_i \right] y(t) = \prod_{j = 1}^{q} \left[ \diff{}{t} - b_i \right] \eta(t),
\end{equation}
where $p, q \in \mathbb{Z}$, $p > q \geq 0$ define the order of the
process; $r_i \in \mathbb{C}$, $\Re r_i < 0$ are the roots of the
characteristic equation of the ODE; $b_j \in \mathbb{C}$,
$\Re b_j < 0$ are corresponding roots of the inverse process; and
$\eta(t)$ is a white-noise Gaussian process with
$\left\langle \eta(t) \right\rangle = 0$ and
$\left\langle \eta(t) \eta(t') \right\rangle = \sigma^2 \delta\left(t
  - t' \right)$.  Such a solution is called a CARMA (Continuous
AutoRegressive Moving Average) process.  The constraint that
$\Re r_i < 0$ and $\Re b_j < 0$ ensures that the linear operator
defining the process and its inverse are invertable.  Under the
assumption that the solution, $y(t)$, is real, roots $r_i$ and $b_j$
must either be real or occur in complex conjugate pairs.  The power
spectrum of the process is
\begin{equation}
  \label{eq:CARMA-PSD}
  P_y(f) = \sigma^2 \left| \frac{\prod_{j=1}^q \left[ 2 \pi i f - b_j \right]}{\prod_{i=1}^p \left[ 2 \pi i f - r_i \right]} \right|^2.
\end{equation}
The power spectrum is a rational function of frequency with poles at
$2\pi i f = r_i$ and zeros at $2\pi i f = b_i$; because of the
invertable constraints, these poles and zeros all occur when
$\Im f \neq 0$.  The autocorrelation function corresponding to the
power spectrum in Eq.\ \eqref{eq:CARMA-PSD} is 
\begin{equation}
  \label{eq:CARMA-ACF}
  \rho_y (\tau) = \left\langle y\left( t \right) y \left( t + \tau \right) \right\rangle = \sum_{i=1}^p A_i^2 e^{r_i \tau},
\end{equation}
where the RMS amplitudes of the $p$ different modes, $A_i$, are
functions of $\sigma$ and the values of the roots $r_i$ and $b_j$.  If
$q = p-1$, then the $A_i$ are independent, while if $q < p-1$ then
there are interdependencies among the $A_i$.  From either the power
spectrum in Eq.\ \eqref{eq:CARMA-PSD} or the autocorrelation function
in Eq.\ \eqref{eq:CARMA-ACF}, it is apparent that a real root $r_i$
represents an exponentially-decaying mode with $r_i$ the rate
constant, while a complex conjugate pair of $r_i$ represent an
oscillatory mode with decay rate $\Re r_i$ and angular frequency
$\left|\Im r_i\right|$.  The former is called an CAR(1) process, an
Ornstein-Uhlenbeck process, or a damped random walk.
\citet{Kelly2014} gives more details on CARMA processes, including
formulae for $A_i$ as a function of $\sigma$, $r_i$, and $b_j$, and
shows how a state-space model of the solution $y(t)$ can be tracked
through a time-series of uncertain observations
$y_k = y\left( t_k \right) + \epsilon_k$, with Gaussian noise
$\epsilon_k$ of zero mean and known (heteroskedastic) variance, using
a Kalman filter to produce a likelihood function
$p\left( y_k \mid \sigma, r_i, b_j \right)$ in
$\mathcal{O}\left( n \right)$ computational time.  Additionally,
\citet{Foreman-Mackey2017} show how a dimensional expansion can be
used to reduce the standard Gaussian process covariance matrix for the
process with autocorrelation given by Eq.\ \eqref{eq:CARMA-ACF} under
observations at times $t_k$ to a banded form that can then be
Cholesky-decomposed in the standard Gaussian process likelihood
function in $\mathcal{O}\left( n \right)$ time.

@@

In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods. 
Aldebaran was observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27. Aldebaran and similar high-luminosity K giants are well known to show photometric variability due to solar-like oscillations \citep{bedding2000}, and hence the continuous, high-precision K2 light curves allow an independent confirmation of our CARMA results.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{insolation.png}
\caption{Evolution of the insolation of Aldebaran~b relative to the incident sunlight on Earth, as a function of the age of Aldebaran on the main sequence. The solid line is the mean of simulations, and the shaded region represents $1\,\sigma$ equivalent deviations. It is likely that Aldebaran~b spent some time on the main sequence with an equilibrium temperature similar to Earth.}
\label{insolation}
\end{figure}
